### 2022-09-20 Prereg_and_manuscript/manuscript.Rmd
manuscript now knits

---
title: "Draft Manuscript"
author: "Emma Menchions"
date: "`r Sys.Date()`"
geometry: margin=1in
header-includes:
    - \usepackage{setspace}
    - \doublespacing
    - \usepackage{lineno}
    - \linenumbers
    - \usepackage{fontspec}         # can only be used with xelatex output
    - \setmainfont{Times New Roman} # can only be used with xelatex output

output:
  pdf_document:
    latex_engine: xelatex

csl: https://www.zotero.org/styles/ecology-letters
bibliography: references.bib
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set
# Installing & Loading packages

#install.packages("groundhog")
library(groundhog)

# installing & loading packages
date <- "2022-09-02"
requiredPackages <-  c("here", "tidyverse","tidyr","Rfast","dplyr")

for (pkg in requiredPackages) {
  if (pkg %in% rownames(installed.packages()) == FALSE)
  {install.packages(pkg)}
  if (pkg %in% rownames(.packages()) == FALSE)
  {groundhog.library(pkg, date)}
}

```

**Title**: Can gradients in annual rainfall & forest type predict the distribution & abundance of 4 Ericaceous shrubs? - BIOL 548T LDP

Emma Menchons^1^\*

1.  Dept. of Botany, University of British Columbia, Vancouver, BC, Canada

\*Corresponding Author

email: emma.menchions\@botany.ubc.ca

**Author Contributions:** EM conceived of the study, conducted the analyses, and wrote the original, and revised drafts of the manuscript.

**Data Availability:** The data and code that support the findings of this study are openly available on Zenodo / GitHub at <https://link_to_archived_release_or_GitHub.com>.

If I was publishing this as an HTML document and wanted the link to be pretty versus human readable, I would use [this format](https://link_to_archived_release_or_GitHub.com) instead

**Conflict of Interest statement**

No conflicts of interest

**Acknowledgements:** We would like to thank a whole bunch of people.

\newpage

# Abstract

1.  Ecologists often have lots of questions about lots of stuff

2.  We evaluated a bunch of things using sophisticated methods and carried out complicated statistical tests

3.  We discovered a bunch of things that we didn't already know but suspected

4.  Our research has greatly advanced out knowledge about stuff and will make a significant contribution to something and someone

**Key-words**: Ecology, Ericaceae, LDP, Plant Ecology, Vegetation Mapping, Vegetation Patterns

\newpage

# Introduction

# Methods

To evaluate the ...

The above demonstrates both "inline" and "display" math formats. If you look at the Source version you will see that it is simply the number of \$ symbols before and after that differentiates between the two. An excellent (and simple) guide on inserting math into your R Markdown documents can be found here:\
<https://rpruim.github.io/s341/S19/from-class/MathinRmd.html>

Statistical analyses were carried out in R 3.4.0 (R Core Team 2017). All code along with the simulation algorithms used are available on Zenodo / GitHub <https://link_to_archived_release_or_GitHub.com>.

# Results

# Discussion

# References

```{=html}
<!-- this littel code block is a comment that won't show up in the knitted 
document. It is a good way to add hidden things and comments to your RMarkdown 
document. For example... -->
```
```{=html}
<!-- this little chunk of code (the ::: stuff below) forces the references to show up here. If you
don't use this then they are put at the end of the document by default -->
```
::: {#refs}
:::

\newpage

# Tables

**Table 1.** Mean body mass of penguins on different islands over time.

Note: for the word version the `kable` does not output correctly. You could try using `flextable` package instead: <https://taehoonh.me/content/post/alternative-to-kable-function-when-knitting-to-ms-word.html>

```{#{r table_01, message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}

# without the "hold_position" the table ends up at the top of the page
```

\newpage

# Figure Captions

**Figure 1**. Pretty coloured dots about penguins

**Figure 2**. Wow, even prettier plot about penguins that shows stuff

\newpage

# Figures

![]()

Figure 1.

\newpage

```{#{r figure_02, echo=FALSE, message=FALSE, warning=FALSE}

```

Figure 2.

\newpage

Figure 3.

```{#{r file="../scripts/figure_03.R"}
# this code chunk calls an external script to generate the plot. If you
# want to load or run code not for a plot you could also use the following.
# Note that it does not seem to work for plots.
# source("../scripts/figure_03.R", local = knitr::knit_global())
```

\newpage

# Appendices

```{=html}
<!-- You might want to look into Bookdown if you want to do anything fancier
with your manuscript. Bookdown allows for automatic Table and Figure numbering
as well as cross referencing. That is you can create links between 'Fig 1' in
your manuscript that links to the actual figure and automatically the
the figure number as you add or remove them from the document.
I add this here because Appendices can require special handling by some 
journals and Bookdown might allow you to do that automagically -->
```