### 2022-09-23 Prereg_and_manuscript/manuscript.Rmd
filling out manuscript introduction

---
title: "Draft Manuscript"
author: "Emma Menchions"
date: "`r Sys.Date()`"
geometry: margin=1in
header-includes:
    - \usepackage{setspace}
    - \doublespacing
    - \usepackage{lineno}
    - \linenumbers
    - \usepackage{fontspec}         # can only be used with xelatex output
    - \setmainfont{Times New Roman} # can only be used with xelatex output

output:
  pdf_document:
    latex_engine: xelatex

csl: https://www.zotero.org/styles/ecology-letters
bibliography: references.bib
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set
# Installing & Loading packages

#install.packages("groundhog")
library(groundhog)

# installing & loading packages
date <- "2022-09-02"
requiredPackages <-  c("ggplot2","here", "tidyverse","tidyr","Rfast","dplyr")

for (pkg in requiredPackages) {
  if (pkg %in% rownames(installed.packages()) == FALSE)
  {install.packages(pkg)}
  if (pkg %in% rownames(.packages()) == FALSE)
  {groundhog.library(pkg, date)}
}

```

**Title**: Can gradients in annual rainfall & forest type predict the distribution & abundance of 4 Ericaceous shrubs? - BIOL 548T LDP

Emma Menchions^1^\*

1.  Dept. of Botany, University of British Columbia, Vancouver, BC, Canada

\*Corresponding Author

email: emma.menchions\@botany.ubc.ca

**Author Contributions:** EM conceived of the study, conducted the analyses, and wrote the original, and revised drafts of the manuscript.

**Data Availability:** The data and code that support the findings of this study are openly available on GitHub at <https://github.com/emench/Menchions-BIOL548T>.

**Conflict of Interest statement**

No conflicts of interest

# Abstract

1.  Ecologists often have lots of questions about lots of stuff

2.  We evaluated a bunch of things using sophisticated methods and carried out complicated statistical tests

3.  We discovered a bunch of things that we didn't already know but suspected

4.  Our research has greatly advanced out knowledge about stuff and will make a significant contribution to something and someone

**Key-words**: Ecology, Ericaceae, LDP, Plant Ecology, Vegetation Mapping, Vegetation Patterns

\newpage

# Introduction

Four species of Ericaceae dominate the understory in the northern boreal forests of Quebec, Canada. Two of these species *Rhododenron groenlandicum* and *Chamaedaphne caluculata* are said to prefer acidic soils and wetter habitats whereas the other dominant species *Kalmia angustifolia*, *Vaccinium angustifolium*, and *Vaccinium myrtillodes* are known to prefer drier, but also acidic soils. Many factors contribute to soil acidity, but a few variables in particular, rainfall and forest type have been observed to drive broad patterns in soil pH in the United States (2). Therefore, I ask, can annual rainfall and forest type (conifer, deciduous, or mixed) predict patterns of abundance for these species in the boreal zone of Northern Quebec between 1985 and 2000?

Acidic soils are often associated with conifer forests and higher rainfall (1,2). Therefore, species with a preference for more acidic soils should be more often found in conifer forest types with higher precipitation. Therefore, I predict that *R. groenlandicum* and *C. calyculata*, will have the highest mean percent cover in areas with conifer forest types and higher annual precipitation. In contrast, I predict that *K. angustifolia* and *Vaccinium sp.*, will have the highest mean percent cover in conifer forests with lower annual precipitation.

Specifically, I predict the following pattern of mean percent cover for *R. groenlandicum* and *C. calyculata*: (H = high precipitation, L = low precipitation, c= conifer forests, d=deciduous forests, m =mixed forest)

From highest to lowest predicted mean percent cover with high rainfall and conifer forests being indicative of higher abundance: H-c, L-c, H-m, H-d, L-m, L-d

For *K. angustifolium* and *Vaccinium sp.* I predict the following pattern, with less rainfall being of greater importance to abundance: L-c, L-m, L-d, H-c, H-m, H-d

# Methods

**Data collection**

Data was retrieved from (3): <https://datadryad.org/stash/dataset/doi:10.5061%2Fdryad.4767v>. According to Thiffault et al. (2016), the data on abundance of the 4 species were collected from 15,339 circular plots across central and northern Quebec along 1.5 -- 2 km long transects of which supported 5-7 plots. Forest type: was determined by visual estimations at the time that the plots were visited by assigning a type corresponding to the two most dominant tree species within the plot.This lead to a total of 15 total forest types. Climate data was collected by identifying the closest meteorological station (51 in the study region) and correcting for differences in latitude, longitude and elevation between the plot and the station using distance weighting and multiple linear regression.

**Data download**

All data download, preparation and statistical analyses were carried out in R 4.2.1 (R Core Team 2022). All scripts used are available on Zenodo / GitHub <https://github.com/emench/Menchions-BIOL548T>. I first retrieved this data from Dryad using a script that automates the download from Dryad.

**Data preparation**

I then prepared the data for analysis by altering the forest type classification and creating a variable indicating whether the plot was in a high or low rainfall area. For the forest type classifications, I used the metadata on forest type codes provided to create 4 categories of forest type: "c" (coniferous, if the two dominant tree species in the plot were both coniferous, n=1282), "d" (deciduous, if the two dominant species were both deciduous, n=45), "m" (mixed, if the two most dominant species contained a conifer and a deciduous species, n=69), and "n" (non-forested, n=26).

I then created a rainfall index. I calculated the median annual usable precipitation (cm) which was 5cm/ year. I used this as a threshold and defined any plot with annual usable precipitation \> 5cm to a high precipitation category (H) and below this threshold to low annual rainfall (L).

Finally, I combined the forest type and rainfall variables into an index (rf index) that described the rainfall category followed by the forest type (e.g. H_c = high rainfall, conifer forest).

**Plotting**

I constructed box plots using using ggplot2::ggplot

# Results

# Discussion

# References

```{=html}
<!-- this littel code block is a comment that won't show up in the knitted 
document. It is a good way to add hidden things and comments to your RMarkdown 
document. For example... -->
```
```{=html}
<!-- this little chunk of code (the ::: stuff below) forces the references to show up here. If you
don't use this then they are put at the end of the document by default -->
```
::: {#refs}
:::

\newpage

# Tables

**Table 1.** Mean body mass of penguins on different islands over time.

Note: for the word version the `kable` does not output correctly. You could try using `flextable` package instead: <https://taehoonh.me/content/post/alternative-to-kable-function-when-knitting-to-ms-word.html>

\`\`\`{#{r table_01, message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}

# without the "hold_position" the table ends up at the top of the page

    \newpage

    # Figure Captions

    **Figure 1**. Pretty coloured dots about penguins

    **Figure 2**. Wow, even prettier plot about penguins that shows stuff

    \newpage

    # Figures

    ![]()

    Figure 1.

    \newpage

    ```{#{r figure_02, echo=FALSE, message=FALSE, warning=FALSE}

Figure 2.

\newpage

Figure 3.

`{#{r file="../scripts/figure_03.R"} # this code chunk calls an external script to generate the plot. If you # want to load or run code not for a plot you could also use the following. # Note that it does not seem to work for plots. # source("../scripts/figure_03.R", local = knitr::knit_global())`

\newpage

# Appendices

```{=html}
<!-- You might want to look into Bookdown if you want to do anything fancier
with your manuscript. Bookdown allows for automatic Table and Figure numbering
as well as cross referencing. That is you can create links between 'Fig 1' in
your manuscript that links to the actual figure and automatically the
the figure number as you add or remove them from the document.
I add this here because Appendices can require special handling by some 
journals and Bookdown might allow you to do that automagically -->
```