### 2022-09-23 Prereg_and_manuscript/manuscript.Rmd
Added code to methods and finished method section

---
title: "Draft Manuscript"
author: "Emma Menchions"
date: "`r Sys.Date()`"
geometry: margin=1in
header-includes:
    - \usepackage{setspace}
    - \doublespacing
    - \usepackage{lineno}
    - \linenumbers
    - \usepackage{fontspec}         # can only be used with xelatex output
    - \setmainfont{Times New Roman} # can only be used with xelatex output

output:
  pdf_document:
    latex_engine: xelatex

csl: https://www.zotero.org/styles/ecology-letters
bibliography: references.bib
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set
# Installing & Loading packages

#install.packages("groundhog")
library(groundhog)

# installing & loading packages
date <- "2022-09-02"
requiredPackages <-  c("ggplot2","here", "tidyverse","tidyr","Rfast","dplyr")

for (pkg in requiredPackages) {
  if (pkg %in% rownames(installed.packages()) == FALSE)
  {install.packages(pkg)}
  if (pkg %in% rownames(.packages()) == FALSE)
  {groundhog.library(pkg, date)}
}

```

**Title**: Can gradients in annual rainfall & forest type predict the distribution & abundance of 4 Ericaceous shrubs? - BIOL 548T LDP

Emma Menchions^1^\*

1.  Dept. of Botany, University of British Columbia, Vancouver, BC, Canada

\*Corresponding Author

email: emma.menchions\@botany.ubc.ca

**Author Contributions:** EM conceived of the study, conducted the analyses, and wrote the original, and revised drafts of the manuscript.

**Data Availability:** The data and code that support the findings of this study are openly available on GitHub at <https://github.com/emench/Menchions-BIOL548T>.

**Conflict of Interest statement**

No conflicts of interest

# Abstract

1.  Ecologists often have lots of questions about lots of stuff

2.  We evaluated a bunch of things using sophisticated methods and carried out complicated statistical tests

3.  We discovered a bunch of things that we didn't already know but suspected

4.  Our research has greatly advanced out knowledge about stuff and will make a significant contribution to something and someone

**Key-words**: Ecology, Ericaceae, LDP, Plant Ecology, Vegetation Mapping, Vegetation Patterns

\newpage

# Introduction

Four species of Ericaceae dominate the understory in the northern boreal forests of Quebec, Canada. Two of these species *Rhododenron groenlandicum* and *Chamaedaphne caluculata* are said to prefer acidic soils and wetter habitats whereas the other dominant species *Kalmia angustifolia*, *Vaccinium angustifolium*, and *Vaccinium myrtillodes* are known to prefer drier, but also acidic soils. Many factors contribute to soil acidity, but a few variables in particular, rainfall and forest type have been observed to drive broad patterns in soil pH in the United States (2). Therefore, I ask, can annual rainfall and forest type (conifer, deciduous, or mixed) predict patterns of abundance for these species in the boreal zone of Northern Quebec between 1985 and 2000?

Acidic soils are often associated with conifer forests and higher rainfall (1,2). Therefore, species with a preference for more acidic soils should be more often found in conifer forest types with higher precipitation. Therefore, I predict that *R. groenlandicum* and *C. calyculata*, will have the highest mean percent cover in areas with conifer forest types and higher annual precipitation. In contrast, I predict that *K. angustifolia* and *Vaccinium sp.*, will have the highest mean percent cover in conifer forests with lower annual precipitation.

Specifically, I predict the following pattern of mean percent cover for *R. groenlandicum* and *C. calyculata*: (H = high precipitation, L = low precipitation, c= conifer forests, d=deciduous forests, m =mixed forest)

From highest to lowest predicted mean percent cover with high rainfall and conifer forests being indicative of higher abundance: H-c, L-c, H-m, H-d, L-m, L-d

For *K. angustifolium* and *Vaccinium sp.* I predict the following pattern, with less rainfall being of greater importance to abundance: L-c, L-m, L-d, H-c, H-m, H-d

# Methods

**Data collection**

Data was retrieved from (3): <https://datadryad.org/stash/dataset/doi:10.5061%2Fdryad.4767v>. According to Thiffault et al. (2016), the data on abundance of the 4 species were collected from 15,339 circular plots across central and northern Quebec along 1.5 -- 2 km long transects of which supported 5-7 plots. Forest type: was determined by visual estimations at the time that the plots were visited by assigning a type corresponding to the two most dominant tree species within the plot.This lead to a total of 15 total forest types. Climate data was collected by identifying the closest meteorological station (51 in the study region) and correcting for differences in latitude, longitude and elevation between the plot and the station using distance weighting and multiple linear regression.

**Data download (script: "part-1_data_download")**

All data download, preparation and statistical analyses were carried out in R 4.2.1 (R Core Team 2022). All scripts used are available on Zenodo / GitHub <https://github.com/emench/Menchions-BIOL548T>. I first retrieved this data from Dryad using a script that automates the download from Dryad.

```{r include=TRUE}
# creating directory to store data
dir.create("data")

## downloading from dryad url ----

data.url <- "https://datadryad.org/stash/downloads/file_stream/41575"
metadata.url <- "https://datadryad.org/stash/downloads/file_stream/41576"

data.dest.file <- paste0("data/","Data_ThriffaultEtAl_EcoEvo_",Sys.Date(),".zip")
metadata.dest.file <- paste0("data/","Metadata_ThriffaultEtAl_EcoEvo_",Sys.Date(),".txt")

unzip(download.file(url = data.url, destfile = data.dest.file),exdir = here::here("data"))
download.file(url= metadata.url, destfile = metadata.dest.file)

unzip(list.files(path = here::here("data"), pattern = "*.zip", full.names = TRUE),exdir = here::here("data"))

```

**Data preparation (script: "part-2_data_preparation")**

I then prepared the data for analysis by altering the forest type classification and creating a variable indicating whether the plot was in a high or low rainfall area. For the forest type classifications, I used the metadata on forest type codes provided to create 4 categories of forest type: "c" (coniferous, if the two dominant tree species in the plot were both coniferous, n=1282), "d" (deciduous, if the two dominant species were both deciduous, n=45), "m" (mixed, if the two most dominant species contained a conifer and a deciduous species, n=69), and "n" (non-forested, n=26).

```{r}
# Loading data tables ----
X_mat <- read.csv(here::here("data","Matrice_X.csv"),
                  sep = ";") # explanatory variables - see metadata
Y_mat <- read.csv(here::here("data","Matrice_Y.csv"),
                  sep = ";") # response variables (mean ericaceous cover, in %)
clim <- read.csv(here::here("data","Climate_value.csv"),
                 sep = ";") # numerical codes used for clim vars in X_mat

# Classifying Forest Type ----
# Identifying two most dominant tree species in each plot and assigning forest type as 
  # conifer (c)
  # deciduous (d)
  # mixed (m)
# looping through each row of X_mat dataframe

for (i in 1:dim(X_mat)[1]){
  most_abun <-  as.numeric(Rfast::nth(as.matrix(X_mat[i,2:14]), k=1, descending = T,index.return = T)) # returning the index of the column with the most abundant tree species for each row
  sec_abun <- as.numeric(Rfast::nth(as.matrix(X_mat[i,2:14]), k=2, descending = T,index.return = T)) # returning index of second most abundant species
  dominant_sp <- c(most_abun, sec_abun)+1 # adding 1 to account for the first column which was left out of vector for previous function
  dominant_sp <- colnames(X_mat)[dominant_sp] # creating character vector containing column names of most dominant species

  if (dominant_sp[1] == "AB" || # if both species conifers
      dominant_sp[1] == "PB" ||
      dominant_sp[1] == "PM"||
      dominant_sp[1] == "PM_AB" ||
      dominant_sp[1] == "TO" &&
      dominant_sp[2] == "AB" ||
      dominant_sp[2] == "PB" ||
      dominant_sp[2] == "PM"||
      dominant_sp[2] == "PM_AM" ||
      dominant_sp[2] == "TO"){
      X_mat$Forest_Type[i] = "c" # forest type conifer

  }else if (dominant_sp[1] == "BA" || # if both species deciduous
            dominant_sp[1] == "BP" ||
            dominant_sp[1] == "PT" &&
            dominant_sp[2] == "BA" ||
            dominant_sp[2] == "BP" ||
            dominant_sp[2] == "PT"){
    X_mat$Forest_Type[i] = "d" # forest type deciduous

  }else if (dominant_sp[1] == "BP_AB" || # if either species is mixed
            dominant_sp[1] == "BP_PM" ||
            dominant_sp[1] == "PT_AB"|| 
            dominant_sp[1] == "PT_PM" ||
            dominant_sp[1] == "BP_AB" ||
            dominant_sp[1] == "BP_PM" ||
            dominant_sp[1] == "PT_AB"|| 
            dominant_sp[1] == "PT_PM"){
    X_mat$Forest_Type[i] = "m"  # forest type mixed

  }else if(dominant_sp[1] == "AB" || # if first species is conifer, second is deciduous 
            dominant_sp[1] == "PB" ||
            dominant_sp[1] == "PM"||
            dominant_sp[1] == "PM_AB" ||
            dominant_sp[1] == "TO" &&
            dominant_sp[2] == "BA" ||
            dominant_sp[2] == "BP" ||
            dominant_sp[2] == "PT"){
    X_mat$Forest_Type[i] = "m" # forest type mixed

  }else if(dominant_sp[1] == "BA" || # if first species is deciduous, second is conifer
            dominant_sp[1] == "BP" ||
            dominant_sp[1] == "PT" &&
            dominant_sp[2] == "AB" ||
            dominant_sp[2] == "PB" ||
            dominant_sp[2] == "PM"||
            dominant_sp[2] == "PM_AB" ||
            dominant_sp[2] == "TO"){
            X_mat$Forest_Type[i] = "m" # forest type mixed
}else{
  X_mat$Forest_Type[i] = "n" # non-forested - other
}
}

# checking the dimensions of the dataframe when we select for specific cateogires
# tells us that the conifer type dominates
# this difference will be normalized later when we take the difference

dim(dplyr::filter(X_mat,Forest_Type=="c")) 
dim(dplyr::filter(X_mat,Forest_Type=="d")) 
dim(dplyr::filter(X_mat,Forest_Type=="m"))
dim(dplyr::filter(X_mat,Forest_Type=="n"))

```

I then created a rainfall index. I calculated the median annual usable precipitation (cm) which was 5cm/ year. I used this as a threshold and defined any plot with annual usable precipitation \> 5cm to a high precipitation category (H) and below this threshold to low annual rainfall (L).

```{r}
# Classifying as high or low precipitation ----
# PRECU is the annual rainfall variable

med <- median(X_mat$PRECU)
# will use the median as the threshold

X_mat <- X_mat %>% 
  mutate(., RainFall = case_when(PRECU < med ~ "L", 
            PRECU >= med ~ "H"))

```

Finally, I combined the forest type and rainfall variables into an index (rf index) that described the rainfall category followed by the forest type (e.g. H_c = high rainfall, conifer forest).

```{r}
# Combining and classifying as forest type - rainfall index
X_mat <- 
  X_mat %>% 
  unite(., rf_index, c("RainFall","Forest_Type"), sep = "_", remove =F, na.rm = FALSE)

# writing altered data as .csv ----
write.csv(X_mat, here::here("data","X_mat_processed.csv"))
```

**Plotting (script: "part-3_data_analysis")**

I constructed box plots displaying percent cover of each species as the response variable and rf (rainfall forest cover index) as the grouping variable using using ggplot2::ggplot. I decided to log-transform the response variable since the data was skewed with outliers making the display and comparison between groups difficult in all of the plots.

```{r}
# loading in prepared data ----
  X_mat <- read.csv(here::here("data","X_mat_processed.csv"))
  X_mat <- X_mat %>% 
    select(CLE, Forest_Type, PRECU, RainFall,rf_index)

  Y_mat <- read.csv(here::here("data","Matrice_Y.csv"),
                    sep = ";") # response variables (mean ericaceous cover, in %)
  clim <- read.csv(here::here("data","Climate_value.csv"),
                   sep = ";") # numerical codes used for clim vars in X_mat

# PLOTTING WITH BOX PLOTS----

  # For Chamaedaphne calyculata
  CAL <- data.frame(Y_mat$V_CAL,X_mat$PRECU, X_mat$rf_index) # creating dataframe with only this species
  colnames(CAL) <- c("V_CAL","PRECU","rf_index") # renaming columns 
  ggplot(data=CAL, aes(x=rf_index, y=V_CAL)) + # plotting only this species with the rainfall-forest type index in box plots
    geom_boxplot()

  ggplot(data=CAL, aes(x=rf_index, y=log(V_CAL))) + # log transforming the y axis for easier pattern detection
    geom_boxplot()  # automatically removes observations with 0 % cover - which is ok

  # For Rhododendron groenlandicum
  RHG <- data.frame(Y_mat$V_RHG,X_mat$PRECU,X_mat$rf_index) # creating dataframe with only this species
  colnames(RHG) <- c("V_RHG","PRECU","rf_index") # renaming columns 
  ggplot(data=RHG, aes(x=rf_index, y=V_RHG)) + # plotting only this species with the rainfall-forest type index in box plots
    geom_boxplot()

  ggplot(data=RHG, aes(x=rf_index, y=log(V_RHG))) + # log transforming the y axis for easier pattern detection
    geom_boxplot()  # automatically removes observations with 0 % cover - which is ok

## For Kalmia angustifolia
  # Plotting: 
  KAA <- data.frame(Y_mat$V_KAA,X_mat$PRECU,X_mat$rf_index) # creating dataframe with only this species
  colnames(KAA) <- c("V_KAA","PRECU","rf_index") # renaming columns

  ggplot(data=KAA, aes(x=rf_index, y=V_KAA)) + # plotting only this species with the rainfall-forest type index in box plots
    geom_boxplot()

  ggplot(data=KAA, aes(x=rf_index, y=log(V_KAA))) + # log transforming the y axis for easier pattern detection
    geom_boxplot()  # automatically removes observations with 0 % cover - which is ok

## For Vaccinium sp. 
    VAAM <- data.frame(Y_mat$V_VAAM,X_mat$PRECU,X_mat$rf_index) # creating dataframe with only this species
    colnames(VAAM) <- c("V_VAAM","PRECU","rf_index") # renaming columns

    ggplot(data=VAAM, aes(x=rf_index, y=V_VAAM)) + # plotting only this species with the rainfall-forest type index in box plots
      geom_boxplot()

    ggplot(data=VAAM, aes(x=rf_index, y=log(V_VAAM))) + # log transforming the y axis for easier pattern detection
      geom_boxplot()  # automatically removes observations with 0 % cover - which is ok

```

I then combined all of the above plots into one single chart for easier comparison.

```{r}
## Plotting altogether in facets 
    # creating new variable to denote species to make table with all species
    CAL <- CAL %>% 
      add_column("species" = "CAL") %>%  # adding column with species names
      rename(cover ="V_CAL") # renaming percent cover column to represent all species

    RHG <- RHG %>% 
      add_column("species" = "RHG") %>% 
      rename(cover ="V_RHG")

    KAA <- KAA %>% 
      add_column("species" = "KAA") %>% 
      rename(cover ="V_KAA")

    VAAM <- VAAM %>% 
      add_column("species" = "VAAM") %>% 
      rename(cover ="V_VAAM")

    plot_cover <- rbind(CAL,RHG, KAA, VAAM) # combining all species by rows into one table

    ## creating ggplot object
    p <- ggplot(data= plot_cover, aes(x=rf_index, y=log(cover))) + # log transforming the y axis for easier pattern detection
      geom_boxplot()+
      xlab("Rainfall-Forest Type Index (RF)") + # changing x axis labe
      ylab("Log (% cover)")+ # changing y axis label
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5), # changing plot themes for angled text and no gridlines
            panel.grid = element_blank())

    ## function to changes names to each plot in facet 
    # from https://stackoverflow.com/questions/3472980/how-to-change-facet-labels
      species_names <- list(
        'CAL'="C. calyculata",
        'KAA'="K. angustifolia",
        'RHG'="R. groenlandicum",
        'VAAM'="Vaccinium sp."
      )

      species_labeller <- function(variable,value){
        return(species_names[value])
      }

    ## creating facet plot with all species plots in one graph 
    p + facet_wrap( ~ species, scales="free",
                    labeller=species_labeller)

    # exporting plot to jpeg
    jpeg(file=here::here("outputs","Cover-RF_Box-Plot.jpeg"))
    p + facet_wrap( ~ species, scales="free",
                    labeller=species_labeller)
    dev.off()
```

**Statistical analysis**

I used a Kruksal-Wallis test to determine if any any groups were significantly different from the rest, or in other words, if at least one rainfall forest cateogry contains a significantly different percent cover of a given species than the other groups. For all species, p \<0.05 indicating we can reject the null hypothesis that percent cover is the same across rainfall forest type categories for all species. To further assess the relationships, I performed a pairwise wilcox test to determine whether percent cover differed significantly from each other in pairs. This was not in the initial pre-registration but is an extended step to further test the hypothesis.

```{r}
 # We want to know is there any significant difference between percent over of plants in different rainfall and forest type conditions

    ## 1) C. calyculata
      # Kruksal-Wallis test to determine if significant difference between means of rf-index groups
      CAL_cover <- plot_cover %>% filter(species=="CAL") # creating daframe with only this species
      kruskal.test(cover ~ rf_index, data = CAL_cover) # performing test - P < 0.05 - yes, indicates significatn differences between groups 
      # use pairwise wilcox test to determine which groups different 
      CAL_wt <- pairwise.wilcox.test(CAL_cover$cover, CAL_cover$rf_index,
                           p.adjust.method = "BH")
      CAL_wt_pframe <- data.frame(CAL_wt[["p.value"]]) # taking p-value matrix and coercing to dataframe
      write.csv(CAL_wt_pframe, here::here("outputs","Ccalyculata_pframe.csv")) # saving dataframe as csv
      write.csv(CAL_wt_pframe>=0.05, here::here("outputs","Ccalyculata_pframe_binary.csv")) # saving data frame indicating p< 0.05 for each paru

      # Creating matrix of difference between medians to pair with the p-value table 
      CAL_meds <- CAL_cover %>% # summarizing by rf group
        group_by(rf_index) %>% 
        summarise(
          median = median(cover, na.rm = TRUE) # calculating median for each group
        )

      CAL_meds_diff <- as.data.frame(outer(CAL_meds$median, CAL_meds$median, FUN= "-")) # creating pairwise difference matrix between medians
      colnames(CAL_meds_diff) <- CAL_meds$rf_index # renaming to have rf index names
      rownames(CAL_meds_diff) <- CAL_meds$rf_index
      write.csv(CAL_meds_diff, here::here("outputs","Ccalyculata_med_differences.csv")) # saving this file as .csv

    ##2) Repeating for K. angustifolia
      # Kruksal-Wallis test to determine if significant difference between means of rf-index groups
      KAA_cover <- plot_cover %>% filter(species=="KAA") # creating daframe with only this species
      kruskal.test(cover ~ rf_index, data = KAA_cover) # performing test - P < 0.05 - yes, indicates significatn differences between groups 
      # use pairwise wilcox test to determine which groups different 
      KAA_wt <- pairwise.wilcox.test(KAA_cover$cover, KAA_cover$rf_index,
                           p.adjust.method = "BH")
      KAA_wt_pframe <- data.frame(KAA_wt[["p.value"]])
      write.csv(KAA_wt_pframe, here::here("outputs","Kangust_pframe.csv"))
      write.csv(KAA_wt_pframe>=0.05, here::here("outputs","Kangust_pframe_binary.csv"))

      # Creating matrix of difference between medians to pair with the p-value table 
      KAA_meds <- KAA_cover %>% 
        group_by(rf_index) %>% 
        summarise(
          median = median(cover, na.rm = TRUE)
        )

      KAA_meds_diff <- as.data.frame(outer(KAA_meds$median, KAA_meds$median, FUN= "-"))
      colnames(KAA_meds_diff) <- KAA_meds$rf_index
      rownames(KAA_meds_diff) <- KAA_meds$rf_index
      write.csv(KAA_meds_diff, here::here("outputs","Kangust_med_differences.csv"))

    ##3) R. groenlandicum
      # Kruksal-Wallis test to determine if significant difference between means of rf-index groups
      RHG_cover <- plot_cover %>% filter(species=="RHG") # creating daframe with only this species
      kruskal.test(cover ~ rf_index, data = RHG_cover) # performing test - P < 0.05 - yes, indicates significatn differences between groups 
      # use pairwise wilcox test to determine which groups different 
      RHG_wt <- pairwise.wilcox.test(RHG_cover$cover, RHG_cover$rf_index,
                           p.adjust.method = "BH")
      RHG_wt_pframe <- data.frame(RHG_wt[["p.value"]])
      write.csv(RHG_wt_pframe, here::here("outputs","Rgroenlandicum_pframe.csv"))
      write.csv(RHG_wt_pframe>=0.05, here::here("outputs","Rgroenlandicum_pframe_binary.csv"))

      # Creating matrix of difference between medians to pair with the p-value table 
      RHG_meds <- RHG_cover %>% 
        group_by(rf_index) %>% 
        summarise(
          median = median(cover, na.rm = TRUE)
        )

      RHG_meds_diff <- as.data.frame(outer(RHG_meds$median, RHG_meds$median, FUN= "-"))
      colnames(RHG_meds_diff) <- RHG_meds$rf_index
      rownames(RHG_meds_diff) <- RHG_meds$rf_index
      write.csv(RHG_meds_diff, here::here("outputs","Rgroenlandicum_med_differences.csv"))

    ##4) Vaccinium sp.
      # Kruksal-Wallis test to determine if significant difference between means of rf-index groups
      VAAM_cover <- plot_cover %>% filter(species=="VAAM") # creating daframe with only this species
      kruskal.test(cover ~ rf_index, data = VAAM_cover) # performing test - P < 0.05 - yes, indicates significatn differences between groups 
      # use pairwise wilcox test to determine which groups different 
      VAAM_wt <- pairwise.wilcox.test(VAAM_cover$cover, VAAM_cover$rf_index,
                           p.adjust.method = "BH")
      VAAM_wt_pframe <- data.frame(VAAM_wt[["p.value"]])
      write.csv(VAAM_wt_pframe, here::here("outputs","Vaccinium_pframe.csv"))
      write.csv(VAAM_wt_pframe>=0.05, here::here("outputs","Vaccinium_pframe_binary.csv"))

      # Creating matrix of difference between medians to pair with the p-value table 
      VAAM_meds <- VAAM_cover %>% 
        group_by(rf_index) %>% 
        summarise(
          median = median(cover, na.rm = TRUE)
        )

      VAAM_meds_diff <- as.data.frame(outer(VAAM_meds$median, VAAM_meds$median, FUN= "-"))
      colnames(VAAM_meds_diff) <- VAAM_meds$rf_index
      rownames(VAAM_meds_diff) <- VAAM_meds$rf_index

```

# Results

\# CAL significant differences (p\< 0.05): H_d vs H_c ; H_m vs H_d ; L_c vs H_d ; L_m vs H_c ; L_m vs L_c

\# KAA significant differences (P\<0.05): H_c vs H_d ; H_c vs H_m ; H_c vs L_c ; H_d vs H_n ; H_m vs H_n ; H_m vs L_m ; H_n vs L_c ; H_n vs L_m

\# RHGsignificant differences (P\<0.05): all types with H_c; H_d vs H_m ; H_d vs H_n ; H_d vs L_c ; L_c vs L_d ; L_c vs L_m ; L_c vs L_n

\# VAAM significant differences (P\<0.05): H_c vs H_d; H_c vs H_m; H_c vs L_c ; H_c vs L_m; H_d vs L_c; H_d vs L_n ; H_m vs L_c; H_m vs L_n

# Discussion

# References

```{=html}
<!-- this littel code block is a comment that won't show up in the knitted 
document. It is a good way to add hidden things and comments to your RMarkdown 
document. For example... -->
```
```{=html}
<!-- this little chunk of code (the ::: stuff below) forces the references to show up here. If you
don't use this then they are put at the end of the document by default -->
```
::: {#refs}
:::

\newpage

# Tables

**Table 1.** Mean body mass of penguins on different islands over time.

Note: for the word version the `kable` does not output correctly. You could try using `flextable` package instead: <https://taehoonh.me/content/post/alternative-to-kable-function-when-knitting-to-ms-word.html>

\`\`\`{#{r table_01, message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}

# without the "hold_position" the table ends up at the top of the page

    \newpage

    # Figure Captions

    **Figure 1**. Pretty coloured dots about penguins

    **Figure 2**. Wow, even prettier plot about penguins that shows stuff

    \newpage

    # Figures

    ![]()

    Figure 1.

    \newpage

    ```{#{r figure_02, echo=FALSE, message=FALSE, warning=FALSE}

Figure 2.

\newpage

Figure 3.

`{#{r file="../scripts/figure_03.R"} # this code chunk calls an external script to generate the plot. If you # want to load or run code not for a plot you could also use the following. # Note that it does not seem to work for plots. # source("../scripts/figure_03.R", local = knitr::knit_global())`

\newpage

# Appendices

```{=html}
<!-- You might want to look into Bookdown if you want to do anything fancier
with your manuscript. Bookdown allows for automatic Table and Figure numbering
as well as cross referencing. That is you can create links between 'Fig 1' in
your manuscript that links to the actual figure and automatically the
the figure number as you add or remove them from the document.
I add this here because Appendices can require special handling by some 
journals and Bookdown might allow you to do that automagically -->
```