### 2022-09-23 Prereg_and_manuscript/manuscript.Rmd
Added resutls and discussion to manuscript

---
title: "Draft Manuscript"
author: "Emma Menchions"
date: "`r Sys.Date()`"
geometry: margin=1in
header-includes:
    - \usepackage{setspace}
    - \doublespacing
    - \usepackage{lineno}
    - \linenumbers
    - \usepackage{fontspec}         # can only be used with xelatex output
    - \setmainfont{Times New Roman} # can only be used with xelatex output

output:
  pdf_document:
    latex_engine: xelatex

csl: https://www.zotero.org/styles/ecology-letters
bibliography: [references.bib,grateful-refs.bib]
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set
# Installing & Loading packages

#install.packages("groundhog")
library(groundhog)

# installing & loading packages
date <- "2022-09-02"
requiredPackages <-  c("ggplot2","here", "tidyverse","tidyr","Rfast","dplyr")

for (pkg in requiredPackages) {
  if (pkg %in% rownames(installed.packages()) == FALSE)
  {install.packages(pkg)}
  if (pkg %in% rownames(.packages()) == FALSE)
  {groundhog.library(pkg, date)}
}

# remotes::install_github("Pakillo/grateful")
# library(grateful)
# get_pkgs_info()
```

**Title**: Can gradients in annual rainfall & forest type predict the distribution & abundance of 4 Ericaceous shrubs? - BIOL 548T LDP

Emma Menchions^1^\*

1.  Dept. of Botany, University of British Columbia, Vancouver, BC, Canada

\*Corresponding Author

email: emma.menchions\@botany.ubc.ca

**Author Contributions:** EM conceived of the study, conducted the analyses, and wrote the original, and revised drafts of the manuscript.

**Data Availability:** The data and code that support the findings of this study are openly available on GitHub at <https://github.com/emench/Menchions-BIOL548T>.

**Conflict of Interest statement**

No conflicts of interest

# Abstract

1.  Ecologists often have lots of questions about lots of stuff

2.  We evaluated a bunch of things using sophisticated methods and carried out complicated statistical tests

3.  We discovered a bunch of things that we didn't already know but suspected

4.  Our research has greatly advanced out knowledge about stuff and will make a significant contribution to something and someone

**Key-words**: Ecology, Ericaceae, LDP, Plant Ecology, Vegetation Mapping, Vegetation Patterns

\newpage

# Introduction

Four species of Ericaceae dominate the understory in the northern boreal forests of Quebec, Canada. Two of these species *Rhododenron groenlandicum* and *Chamaedaphne caluculata* are said to prefer acidic soils and wetter habitats whereas the other dominant species *Kalmia angustifolia*, *Vaccinium angustifolium*, and *Vaccinium myrtillodes* are known to prefer drier, but also acidic soils. Many factors contribute to soil acidity, but a few variables in particular, rainfall and forest type have been observed to drive broad patterns in soil pH in the United States (2). Therefore, I ask, can annual rainfall and forest type (conifer, deciduous, or mixed) predict patterns of abundance for these species in the boreal zone of Northern Quebec between 1985 and 2000?

Acidic soils are often associated with conifer forests and higher rainfall (1,2). Therefore, species with a preference for more acidic soils should be more often found in conifer forest types with higher precipitation. Therefore, I predict that *R. groenlandicum* and *C. calyculata*, will have the highest mean percent cover in areas with conifer forest types and higher annual precipitation. In contrast, I predict that *K. angustifolia* and *Vaccinium sp.*, will have the highest mean percent cover in conifer forests with lower annual precipitation.

Specifically, I predict the following pattern of mean percent cover for *R. groenlandicum* and *C. calyculata*: (H = high precipitation, L = low precipitation, c= conifer forests, d=deciduous forests, m =mixed forest)

From highest to lowest predicted mean percent cover with high rainfall and conifer forests being indicative of higher abundance: H-c, L-c, H-m, H-d, L-m, L-d

For *K. angustifolium* and *Vaccinium sp.* I predict the following pattern, with less rainfall being of greater importance to abundance: L-c, L-m, L-d, H-c, H-m, H-d

# Methods

**Data collection**

Data was retrieved from (3): <https://datadryad.org/stash/dataset/doi:10.5061%2Fdryad.4767v>. According to Thiffault et al. (2016), the data on abundance of the 4 species were collected from 15,339 circular plots across central and northern Quebec along 1.5 -- 2 km long transects of which supported 5-7 plots. Forest type: was determined by visual estimations at the time that the plots were visited by assigning a type corresponding to the two most dominant tree species within the plot.This lead to a total of 15 total forest types. Climate data was collected by identifying the closest meteorological station (51 in the study region) and correcting for differences in latitude, longitude and elevation between the plot and the station using distance weighting and multiple linear regression.

**Data download (script: "part-1_data_download")**

All data download, preparation and statistical analyses were carried out in R 4.2.1 (R Core Team 2022). All scripts used are available on Zenodo / GitHub <https://github.com/emench/Menchions-BIOL548T>. I first retrieved this data from Dryad using a script that automates the download from Dryad.

```{r eval=FALSE, warning=FALSE, include=FALSE}
# creating directory to store data
dir.create("data")

## downloading from dryad url ----

data.url <- "https://datadryad.org/stash/downloads/file_stream/41575"
metadata.url <- "https://datadryad.org/stash/downloads/file_stream/41576"

data.dest.file <- paste0("data/","Data_ThriffaultEtAl_EcoEvo_",Sys.Date(),".zip")
metadata.dest.file <- paste0("data/","Metadata_ThriffaultEtAl_EcoEvo_",Sys.Date(),".txt")

unzip(download.file(url = data.url, destfile = data.dest.file),exdir = here::here("data"))
download.file(url= metadata.url, destfile = metadata.dest.file)

unzip(list.files(path = here::here("data"), pattern = "*.zip", full.names = TRUE),exdir = here::here("data"))

```

**Data preparation (script: "part-2_data_preparation")**

I then prepared the data for analysis by altering the forest type classification and creating a variable indicating whether the plot was in a high or low rainfall area. For the forest type classifications, I used the metadata on forest type codes provided to create 4 categories of forest type: "c" (coniferous, if the two dominant tree species in the plot were both coniferous, n=1282), "d" (deciduous, if the two dominant species were both deciduous, n=45), "m" (mixed, if the two most dominant species contained a conifer and a deciduous species, n=69), and "n" (non-forested, n=26).

```{r, include=TRUE,echo=TRUE}
# Loading data tables ----
X_mat <- read.csv(here::here("data","Matrice_X.csv"),
                  sep = ";") # explanatory variables - see metadata
Y_mat <- read.csv(here::here("data","Matrice_Y.csv"),
                  sep = ";") # response variables (mean ericaceous cover, in %)
clim <- read.csv(here::here("data","Climate_value.csv"),
                 sep = ";") # numerical codes used for clim vars in X_mat

# Classifying Forest Type ----
# Identifying two most dominant tree species in each plot and assigning forest type as 
  # conifer (c)
  # deciduous (d)
  # mixed (m)
# looping through each row of X_mat dataframe

for (i in 1:dim(X_mat)[1]){
  most_abun <-  as.numeric(Rfast::nth(as.matrix(X_mat[i,2:14]), 
                                      k=1, descending = T,index.return = T)) 
  # returning the index of the column with the most abundant tree species for each row
  sec_abun <- as.numeric(Rfast::nth(as.matrix(X_mat[i,2:14]), 
                                    k=2, descending = T,index.return = T)) 
  # returning index of second most abundant species
  dominant_sp <- c(most_abun, sec_abun)+1 # adding 1 to account for the 
  # first column which was left out of vector for previous function
  dominant_sp <- colnames(X_mat)[dominant_sp] # creating character vector containing 
  # column names of most dominant species

  if (dominant_sp[1] == "AB" || # if both species conifers
      dominant_sp[1] == "PB" ||
      dominant_sp[1] == "PM"||
      dominant_sp[1] == "PM_AB" ||
      dominant_sp[1] == "TO" &&
      dominant_sp[2] == "AB" ||
      dominant_sp[2] == "PB" ||
      dominant_sp[2] == "PM"||
      dominant_sp[2] == "PM_AM" ||
      dominant_sp[2] == "TO"){
      X_mat$Forest_Type[i] = "c" # forest type conifer

  }else if (dominant_sp[1] == "BA" || # if both species deciduous
            dominant_sp[1] == "BP" ||
            dominant_sp[1] == "PT" &&
            dominant_sp[2] == "BA" ||
            dominant_sp[2] == "BP" ||
            dominant_sp[2] == "PT"){
    X_mat$Forest_Type[i] = "d" # forest type deciduous

  }else if (dominant_sp[1] == "BP_AB" || # if either species is mixed
            dominant_sp[1] == "BP_PM" ||
            dominant_sp[1] == "PT_AB"|| 
            dominant_sp[1] == "PT_PM" ||
            dominant_sp[1] == "BP_AB" ||
            dominant_sp[1] == "BP_PM" ||
            dominant_sp[1] == "PT_AB"|| 
            dominant_sp[1] == "PT_PM"){
    X_mat$Forest_Type[i] = "m"  # forest type mixed

  }else if(dominant_sp[1] == "AB" || # if first species is conifer, second is deciduous 
            dominant_sp[1] == "PB" ||
            dominant_sp[1] == "PM"||
            dominant_sp[1] == "PM_AB" ||
            dominant_sp[1] == "TO" &&
            dominant_sp[2] == "BA" ||
            dominant_sp[2] == "BP" ||
            dominant_sp[2] == "PT"){
    X_mat$Forest_Type[i] = "m" # forest type mixed

  }else if(dominant_sp[1] == "BA" || # if first species is deciduous, second is conifer
            dominant_sp[1] == "BP" ||
            dominant_sp[1] == "PT" &&
            dominant_sp[2] == "AB" ||
            dominant_sp[2] == "PB" ||
            dominant_sp[2] == "PM"||
            dominant_sp[2] == "PM_AB" ||
            dominant_sp[2] == "TO"){
            X_mat$Forest_Type[i] = "m" # forest type mixed
}else{
  X_mat$Forest_Type[i] = "n" # non-forested - other
}
}

# checking the dimensions of the dataframe when we select for specific cateogires
# tells us that the conifer type dominates
# this difference will be normalized later when we take the difference

dim(dplyr::filter(X_mat,Forest_Type=="c")) 
dim(dplyr::filter(X_mat,Forest_Type=="d")) 
dim(dplyr::filter(X_mat,Forest_Type=="m"))
dim(dplyr::filter(X_mat,Forest_Type=="n"))

```

I then created a rainfall index. I calculated the median annual usable precipitation (cm) which was 5cm/ year. I used this as a threshold and defined any plot with annual usable precipitation \> 5cm to a high precipitation category (H) and below this threshold to low annual rainfall (L).

```{r,include=TRUE,echo=TRUE}
# Classifying as high or low precipitation ----
# PRECU is the annual rainfall variable

med <- median(X_mat$PRECU)
# will use the median as the threshold

X_mat <- X_mat %>% 
  mutate(., RainFall = case_when(PRECU < med ~ "L", 
            PRECU >= med ~ "H"))

```

Finally, I combined the forest type and rainfall variables into an index (rf index) that described the rainfall category followed by the forest type (e.g. H_c = high rainfall, conifer forest).

```{r,include=TRUE,echo=TRUE}
# Combining and classifying as forest type - rainfall index
X_mat <- 
  X_mat %>% 
  unite(., rf_index, c("RainFall","Forest_Type"), sep = "_", remove =F, na.rm = FALSE)

# writing altered data as .csv ----
write.csv(X_mat, here::here("data","X_mat_processed.csv"))
```

**Plotting (script: "part-3_data_analysis")**

I constructed box plots displaying percent cover of each species as the response variable and rf (rainfall forest cover index) as the grouping variable using using ggplot2::ggplot. I decided to log-transform the response variable since the data was skewed with outliers making the display and comparison between groups difficult in all of the plots.

```{r, include=TRUE,echo=TRUE}
# loading in prepared data ----
  X_mat <- read.csv(here::here("data","X_mat_processed.csv"))
  X_mat <- X_mat %>% 
    select(CLE, Forest_Type, PRECU, RainFall,rf_index)

  Y_mat <- read.csv(here::here("data","Matrice_Y.csv"),
                    sep = ";") # response variables (mean ericaceous cover, in %)
  clim <- read.csv(here::here("data","Climate_value.csv"),
                   sep = ";") # numerical codes used for clim vars in X_mat

# PLOTTING WITH BOX PLOTS----

  # For Chamaedaphne calyculata
  # creating dataframe with only this species
  CAL <- data.frame(Y_mat$V_CAL,X_mat$PRECU, X_mat$rf_index) 
  colnames(CAL) <- c("V_CAL","PRECU","rf_index") # renaming columns 
  ggplot(data=CAL, aes(x=rf_index, y=V_CAL)) + 
    geom_boxplot() # plotting only this species with the rainfall-forest
    # type index in box plots

  CAL_plot <- ggplot(data=CAL, aes(x=rf_index, y=log(V_CAL))) + # log transforming the y axis 
    geom_boxplot()  # automatically removes observations with 0 % cover - which is ok

  # Repeating for Rhododendron groenlandicum
  RHG <- data.frame(Y_mat$V_RHG,X_mat$PRECU,X_mat$rf_index) 
  colnames(RHG) <- c("V_RHG","PRECU","rf_index") 
  RHG_plot <- ggplot(data=RHG, aes(x=rf_index, y=V_RHG)) + 
    geom_boxplot()

  RHG_plot <- ggplot(data=RHG, aes(x=rf_index, y=log(V_RHG))) + # log transforming the y axis 
    geom_boxplot()

## Repeating for Kalmia angustifolia
  # Plotting: 
  KAA <- data.frame(Y_mat$V_KAA,X_mat$PRECU,X_mat$rf_index) 
  colnames(KAA) <- c("V_KAA","PRECU","rf_index") 
  KAA_plot <- ggplot(data=KAA, aes(x=rf_index, y=V_KAA)) + 
    geom_boxplot()

  KAA_plot <- ggplot(data=KAA, aes(x=rf_index, y=log(V_KAA))) + # log transforming the y axis 
    geom_boxplot()

## Repeating for Vaccinium sp. 
    VAAM <- data.frame(Y_mat$V_VAAM,X_mat$PRECU,X_mat$rf_index) 
    colnames(VAAM) <- c("V_VAAM","PRECU","rf_index")

   VAAM_plot <-ggplot(data=VAAM, aes(x=rf_index, y=V_VAAM)) +
      geom_boxplot()

    VAAM_plot <- ggplot(data=VAAM, aes(x=rf_index, y=log(V_VAAM))) + # log transforming the y axis 
      geom_boxplot()

```

I then combined all of the above plots into one single chart for easier comparison.

```{r,echo=FALSE, include=FALSE}
## Plotting altogether in facets 
    # creating new variable to denote species to make table with all species
    CAL <- CAL %>% 
      add_column("species" = "CAL") %>%  # adding column with species names
      rename(cover ="V_CAL") # renaming percent cover column to represent all species

    RHG <- RHG %>% 
      add_column("species" = "RHG") %>% 
      rename(cover ="V_RHG")

    KAA <- KAA %>% 
      add_column("species" = "KAA") %>% 
      rename(cover ="V_KAA")

    VAAM <- VAAM %>% 
      add_column("species" = "VAAM") %>% 
      rename(cover ="V_VAAM")

    plot_cover <- rbind(CAL,RHG, KAA, VAAM) # combining all species by rows into table

    ## creating ggplot object
    p <- ggplot(data= plot_cover, aes(x=rf_index, y=log(cover))) + # log transforming 
      geom_boxplot()+
      xlab("Rainfall-Forest Type Index (RF)") + # changing x axis labe
      ylab("Log (% cover)")+ # changing y axis label
      # changing plot themes for angled text and no gridlines
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5), 
            panel.grid = element_blank())

    ## function to changes names to each plot in facet 
    # from https://stackoverflow.com/questions/3472980/how-to-change-facet-labels
      species_names <- list(
        'CAL'="C. calyculata",
        'KAA'="K. angustifolia",
        'RHG'="R. groenlandicum",
        'VAAM'="Vaccinium sp."
      )

      species_labeller <- function(variable,value){
        return(species_names[value])
      }

    ## creating facet plot with all species plots in one graph 
    p + facet_wrap( ~ species, scales="free",
                    labeller=species_labeller)

    # exporting plot to jpeg
    jpeg(file=here::here("outputs","Cover-RF_Box-Plot.jpeg"))
    p + facet_wrap( ~ species, scales="free",
                    labeller=species_labeller)
    dev.off()
```

**Statistical analysis**

I used a Kruksal-Wallis test to determine if any any groups were significantly different from the rest, or in other words, if at least one rainfall forest cateogry contains a significantly different percent cover of a given species than the other groups. For all species, p \<0.05 indicating we can reject the null hypothesis that percent cover is the same across rainfall forest type categories for all species. To further assess the relationships, I performed a pairwise wilcox test to determine whether percent cover differed significantly from each other in pairs. This was not in the initial pre-registration but is an extended step to further test the hypothesis.

```{r,echo=FALSE, include=FALSE}
 # We want to know is there any significant difference between percent over of plants in different rainfall and forest type conditions

    ## 1) C. calyculata
      # Kruksal-Wallis test to determine if significant difference between means of rf-index groups
      CAL_cover <- plot_cover %>% filter(species=="CAL") # creating daframe with only this species
      kruskal.test(cover ~ rf_index, data = CAL_cover) # performing test - P < 0.05 - yes, indicates significatn differences between groups 
      # use pairwise wilcox test to determine which groups different 
      CAL_wt <- pairwise.wilcox.test(CAL_cover$cover, CAL_cover$rf_index,
                           p.adjust.method = "BH")
      CAL_wt_pframe <- data.frame(CAL_wt[["p.value"]]) # taking p-value matrix and coercing to dataframe
      write.csv(CAL_wt_pframe, here::here("outputs","Ccalyculata_pframe.csv")) # saving dataframe as csv
      write.csv(CAL_wt_pframe>=0.05, here::here("outputs","Ccalyculata_pframe_binary.csv")) # saving data frame indicating p< 0.05 for each paru

      # Creating matrix of difference between medians to pair with the p-value table 
      CAL_meds <- CAL_cover %>% # summarizing by rf group
        group_by(rf_index) %>% 
        summarise(
          median = median(cover, na.rm = TRUE) # calculating median for each group
        )

      CAL_meds_diff <- as.data.frame(outer(CAL_meds$median, CAL_meds$median, FUN= "-")) # creating pairwise difference matrix between medians
      colnames(CAL_meds_diff) <- CAL_meds$rf_index # renaming to have rf index names
      rownames(CAL_meds_diff) <- CAL_meds$rf_index
      write.csv(CAL_meds_diff, here::here("outputs","Ccalyculata_med_differences.csv")) # saving this file as .csv

    ##2) Repeating for K. angustifolia
      # Kruksal-Wallis test to determine if significant difference between means of rf-index groups
      KAA_cover <- plot_cover %>% filter(species=="KAA") # creating daframe with only this species
      kruskal.test(cover ~ rf_index, data = KAA_cover) # performing test - P < 0.05 - yes, indicates significatn differences between groups 
      # use pairwise wilcox test to determine which groups different 
      KAA_wt <- pairwise.wilcox.test(KAA_cover$cover, KAA_cover$rf_index,
                           p.adjust.method = "BH")
      KAA_wt_pframe <- data.frame(KAA_wt[["p.value"]])
      write.csv(KAA_wt_pframe, here::here("outputs","Kangust_pframe.csv"))
      write.csv(KAA_wt_pframe>=0.05, here::here("outputs","Kangust_pframe_binary.csv"))

      # Creating matrix of difference between medians to pair with the p-value table 
      KAA_meds <- KAA_cover %>% 
        group_by(rf_index) %>% 
        summarise(
          median = median(cover, na.rm = TRUE)
        )

      KAA_meds_diff <- as.data.frame(outer(KAA_meds$median, KAA_meds$median, FUN= "-"))
      colnames(KAA_meds_diff) <- KAA_meds$rf_index
      rownames(KAA_meds_diff) <- KAA_meds$rf_index
      write.csv(KAA_meds_diff, here::here("outputs","Kangust_med_differences.csv"))

    ##3) R. groenlandicum
      # Kruksal-Wallis test to determine if significant difference between means of rf-index groups
      RHG_cover <- plot_cover %>% filter(species=="RHG") # creating daframe with only this species
      kruskal.test(cover ~ rf_index, data = RHG_cover) # performing test - P < 0.05 - yes, indicates significatn differences between groups 
      # use pairwise wilcox test to determine which groups different 
      RHG_wt <- pairwise.wilcox.test(RHG_cover$cover, RHG_cover$rf_index,
                           p.adjust.method = "BH")
      RHG_wt_pframe <- data.frame(RHG_wt[["p.value"]])
      write.csv(RHG_wt_pframe, here::here("outputs","Rgroenlandicum_pframe.csv"))
      write.csv(RHG_wt_pframe>=0.05, here::here("outputs","Rgroenlandicum_pframe_binary.csv"))

      # Creating matrix of difference between medians to pair with the p-value table 
      RHG_meds <- RHG_cover %>% 
        group_by(rf_index) %>% 
        summarise(
          median = median(cover, na.rm = TRUE)
        )

      RHG_meds_diff <- as.data.frame(outer(RHG_meds$median, RHG_meds$median, FUN= "-"))
      colnames(RHG_meds_diff) <- RHG_meds$rf_index
      rownames(RHG_meds_diff) <- RHG_meds$rf_index
      write.csv(RHG_meds_diff, here::here("outputs","Rgroenlandicum_med_differences.csv"))

    ##4) Vaccinium sp.
      # Kruksal-Wallis test to determine if significant difference between means of rf-index groups
      VAAM_cover <- plot_cover %>% filter(species=="VAAM") # creating daframe with only this species
      kruskal.test(cover ~ rf_index, data = VAAM_cover) # performing test - P < 0.05 - yes, indicates significatn differences between groups 
      # use pairwise wilcox test to determine which groups different 
      VAAM_wt <- pairwise.wilcox.test(VAAM_cover$cover, VAAM_cover$rf_index,
                           p.adjust.method = "BH")
      VAAM_wt_pframe <- data.frame(VAAM_wt[["p.value"]])
      write.csv(VAAM_wt_pframe, here::here("outputs","Vaccinium_pframe.csv"))
      write.csv(VAAM_wt_pframe>=0.05, here::here("outputs","Vaccinium_pframe_binary.csv"))

      # Creating matrix of difference between medians to pair with the p-value table 
      VAAM_meds <- VAAM_cover %>% 
        group_by(rf_index) %>% 
        summarise(
          median = median(cover, na.rm = TRUE)
        )

      VAAM_meds_diff <- as.data.frame(outer(VAAM_meds$median, VAAM_meds$median, FUN= "-"))
      colnames(VAAM_meds_diff) <- VAAM_meds$rf_index
      rownames(VAAM_meds_diff) <- VAAM_meds$rf_index

```

# Results & Discussion

```{r}
p + facet_wrap( ~ species, scales="free",
                    labeller=species_labeller)
```

**Figure 1.** Box plots of *C. calyculata*, *K. angustifolia*, *R. groenlandicum*, and *Vaccinium sp*. percent cover in plots that fall in each of the rainfall-forest type categories. The thick horizontal lines represent the median for each category. The y axis was log-transformed for display purposes.

**Table 1.** Mean % cover for *C. calyculata* in different rainfall forest type categories.

```{r}
CAL_cover %>% group_by(rf_index) %>% summarise(median = median(cover, na.rm=T)) %>% arrange(desc(median))
```

For *C. calyculata,* the only rainfall forest type category that had median percent cover above 0 was high rainfall coniferous forests. This supports the prediction that this species would have a greater preference for this category possibly due to the higher acidity and moisture provided.

**Table 2.** Mean % cover for *R. groenlandicum* in different rainfall forest type categories.

```{r}
RHG_cover %>% group_by(rf_index) %>% summarise(median = median(cover, na.rm=T)) %>% arrange(desc(median))
```

For *R. groenlandicum*, the order of my predictions also matched those observed. H_c had significantly higher percent cover than any other rf type (p\<0.05), followed by L-c, H_m, L_m H_d and L_d. This may indicate that *R. groenlandicum* indeed prefers more acidic soils created by higher rainfall, but mroe importantly coniferous forests. However, this is confounded with other factors that may drive its preference for coniferous habitat such as other soil properties associated with conifer needles and light availability.

**Table 3.** Mean % cover for *K. angustifolium* in different rainfall forest type categories.

```{r}
KAA_cover %>% group_by(rf_index) %>% summarise(median = median(cover, na.rm=T)) %>% arrange(desc(median))
```

For *K. angustifolium* the patterns in abundance that I predicted (L-c, L-m, L-d, H-c, H-m, H-d) were not supported by those observed. The category with the highest median percent cover was non-forested area however this was not significantly different from abundance in high rainfall coniferous forests (p\>0.05). This may indicate that light-availability is a more important determinant of its distribution. Additionally, non-forested plots may often be boggy areas with acidic soils from sphagnum growth, so perhaps this species' distribution is highly driven by soil acidity, so much so that the acidity produced by conifer needles and high rainfall is not sufficient.

```{r}
VAAM_cover %>% group_by(rf_index) %>% summarise(median = median(cover, na.rm=T)) %>% arrange(desc(median))
```

Finally, for *Vaccinium sp.*

For *K. angustifolium* and *Vaccinium sp.* I predict the following pattern, with less rainfall being of greater importance to abundance: L-c, L-m, L-d, H-c, H-m, H-d

\# KAA significant differences (P\<0.05): H_c vs H_d ; H_c vs H_m ; H_c vs L_c ; H_d vs H_n ; H_m vs H_n ; H_m vs L_m ; H_n vs L_c ; H_n vs L_m

\# VAAM significant differences (P\<0.05): H_c vs H_d; H_c vs H_m; H_c vs L_c ; H_c vs L_m; H_d vs L_c; H_d vs L_n ; H_m vs L_c; H_m vs L_n

Analysis to test if these species distributions are driven partly by soil acidity could ideally be performed using direct measures of soil acidity at the plots rather than the proxy variables used here. Upon future visits to the sites, these measurements could be taken to construct models and compare preferences between species to better predict what drives their distribution patterns and regenerative ability in logging and fire disturbed habitats.

# References

```{=html}
<!-- this littel code block is a comment that won't show up in the knitted 
document. It is a good way to add hidden things and comments to your RMarkdown 
document. For example... -->
```
```{=html}
<!-- this little chunk of code (the ::: stuff below) forces the references to show up here. If you
don't use this then they are put at the end of the document by default -->
```
::: {#refs}
:::

```{r results = 'asis'}
grateful::cite_packages(output = "paragraph")
```

\newpage

# Appendix

**Table A1.**